---
title: "kallisto-tximport-DESeq2"
author: "Serdar Turkarslan"
date: "09/30/2020"
output: html_document
---

```{r setup, include=FALSE}
library('tximport')
library('rhdf5')
library("DESeq2")
library('DT')
library('tictoc')
library('apeglm')
library('viridis')
library('tidyverse')
library('Cairo')
source('code/import_kallisto2DESeq.R')
source('code/create_conditions.R')
source('code/deg_analysis.R')
source('code/volcano_plot.R')
source('code/create_DEG_matrix.R')
source('code/plot_pca.R')
source('code/plot_topn_heatmap.R')
Sys.setenv("DISPLAY"=":0.0")

```

## Load kallisto counts.
<p>Get list of abundance files from kallisto results and load into DESEq</p>
```{r echo=TRUE, message=FALSE, warning=TRUE,results='hide'}
if(fs::is_file("data/deseq_data.RDS")){
   cat("\t DESEq2 result data exists, loading.. \n")
   deseq_data <- readRDS('data/deseq_data.RDS')
 }else{
   cat("\t DESEq2 result data does not exist, starting analysis.. \n")
   deseq_data <- import_kallisto2DESeq()
   saveRDS(deseq_data, file = "data/deseq_data.RDS")
 }
datatable(deseq_data$meta_data, caption = "Meta-data information")
```

## Create comparison conditions {.tabset}
Create conditions for differential expression analysis {.tabset}
```{r echo=TRUE, message=FALSE, warning=FALSE,results='hide'}
conditions <- create_conditions(meta_data= deseq_data$meta_data)
```


### DEG analysis for Spis
```{r echo=TRUE, message=FALSE,warning=FALSE,results='hide'}
cat("Started DEG analysis...\n")
org = "Spis"
for(condition in conditions){
   deg.file <- paste("output/", org, "/DEG_tables_svalues/",condition,".txt", sep="")
   if(fs::is_file(deg.file)){
      cat("\t DEG analysis for ", condition, "exists, skipping \n")
   }else{
      cat("\t DEG analysis for ", condition, "does NOT exists, continuing \n")
      deg_analysis(condition=condition, dds=deseq_data$dds.Spis, lfc=2, write2file=T, org = org)
   }
}
```
<p>DEG Results for all conditions are [here](`r paste("output/", org, "/DEG_tables_svalues/", sep="")`)</p>


#### Volcano plots
```{r echo=TRUE,message=FALSE,warning=FALSE,results='hide'}
cat("Plotting volcano plots.\n")
org = "Spis"
rerun = FALSE
for(condition in conditions){
   pdf.file <- paste("output/",org,"/volcano_plots_svalues/",condition,"_volcano.pdf", sep="")
   if(fs::is_file(pdf.file) & rerun==FALSE){
      cat("\t Volcano plot for ", condition, "exists, skipping \n")
   }else{
      cat("\t Volcano plot for ", condition, "does NOT exists, continuing \n")
      pdf(file=pdf.file)
      volcano_plot(condition = condition, org="Spis", lfc=2, svalue=0.005)
      dev.off()
   }
}
```
<p>Volcano plots for all conditions are [here](`r paste("output/", org, "/volcano_plots_svalues/", sep="")`)</p>


#### DEG Matrix
```{r echo=TRUE,dev=c('png'),results='hide'}
##### 4. create DEG matrix and write results into a file ####
org="Spis"

DEG.matrix <- create_DEG_matrix(org = org)
write.table(DEG.matrix$matrix, file=paste("output/",org,"/DEG_matrix.txt", sep=""), sep="\t")
write.table(DEG.matrix$updown, file=paste("output/",org,"/DEG_matrix_up_down.txt", sep=""), sep="\t")
pheatmap(mat=DEG.matrix$matrix,
         cluster_rows = F, 
         cluster_cols = F, 
         display_numbers = T,
         number_format = "%.0f", 
         main = paste(org, " DEG Matrix", sep = "")
         )
dev.print(pdf, paste("output/",org,"/DEG_matrix_heatmap.pdf", sep=""))
```

#### Heatmap of top 50 DEGs
```{r echo=TRUE,dev=c('png')}
org="Spis"
results_dir= paste("output/", org,"/summary_plots/", sep = "")
plot_topn_heatmap(org = "Spis",n = 50,dds=deseq_data$dds.Spis)
dev.print(pdf, paste(results_dir,"top50_DEG_heatmap.pdf",sep = ""))
```

#### PCA for all samples
```{r echo=TRUE,dev=c('png')}
org="Spis"
results_dir= paste("output/", org,"/summary_plots/", sep = "")
plot_pca(org = "Spis",vsd = deseq_data$vsd.Spis)
dev.print(pdf, paste(results_dir,"PCA_plot.pdf",sep = ""))
```

#### {-}

### DEG analysis for Smic
```{r echo=TRUE, message=FALSE,warning=FALSE,results='hide'}
cat("Started DEG analysis...\n")
org = "Smic"
for(condition in conditions){
   deg.file <- paste("output/", org, "/DEG_tables_svalues/",condition,".txt", sep="")
   if(fs::is_file(deg.file)){
      cat("\t DEG analysis for ", condition, "exists, skipping \n")
   }else{
      cat("\t DEG analysis for ", condition, "does NOT exists, continuing \n")
      deg_analysis(condition=condition, dds=deseq_data$dds.Smic, lfc=2, write2file=T, org = org)
   }
}
```
<p>DEG Results for all conditions are [here](`r paste("output/", org, "/DEG_tables_svalues/", sep="")`)</p>


#### Volcano plots
```{r echo=TRUE,message=FALSE,warning=FALSE,results='hide'}
cat("Plotting volcano plots.\n")
org = "Smic"
rerun = FALSE
for(condition in conditions){
   pdf.file <- paste("output/",org,"/volcano_plots_svalues/",condition,"_volcano.pdf", sep="")
   if(fs::is_file(pdf.file) & rerun==FALSE){
      cat("\t Volcano plot for ", condition, "exists, skipping \n")
   }else{
      cat("\t Volcano plot for ", condition, "does NOT exists, continuing \n")
      pdf(file=pdf.file)
      volcano_plot(condition = condition, org=org, lfc=2, svalue=0.005)
      dev.off()
   }
}
```
<p>Volcano plots for all conditions are [here](`r paste("output/", org, "/volcano_plots_svalues/", sep="")`)</p>


#### DEG Matrix
```{r echo=TRUE,dev=c('png'),results='hide'}
##### 4. create DEG matrix and write results into a file #####
org="Smic"
DEG.matrix <- create_DEG_matrix(org = org)
write.table(DEG.matrix$matrix, file=paste("output/",org,"/DEG_matrix.txt", sep=""), sep="\t")
write.table(DEG.matrix$updown, file=paste("output/",org,"/DEG_matrix_up_down.txt", sep=""), sep="\t")
pheatmap(mat=DEG.matrix$matrix,
         cluster_rows = F, 
         cluster_cols = F, 
         display_numbers = T,
         number_format = "%.0f", 
         main = paste(org, " DEG Matrix", sep = "")
         )
dev.print(pdf, paste("output/",org,"/DEG_matrix_heatmap.pdf", sep=""))
```

#### Heatmap of top 50 DEGs
```{r echo=TRUE,dev=c('png')}
org="Smic"
results_dir= paste("output/", org,"/summary_plots/", sep = "")
plot_topn_heatmap(org = "Smic",n = 50,dds=deseq_data$dds.Smic)
dev.print(pdf, paste(results_dir,"top50_DEG_heatmap.pdf",sep = ""))
```

#### PCA for all samples
```{r echo=TRUE,dev=c('png')}
org="Smic"
results_dir= paste("output/", org,"/summary_plots/", sep = "")
plot_pca(org = org,vsd = deseq_data$vsd.Smic)
dev.print(pdf, paste(results_dir,"PCA_plot.pdf",sep = ""))
```

#### {-}

